Parsing Postfix log files

### 2008-10-28 doc/logparser-introduction.tex
First pass through corrections for sections 1-3.

Most mail server administrators will have performed some basic processing
of the log files produced by their mail server at one time or another,
whether it was to debug a problem, explain to a user why their mail is
being rejected, or check if new anti-spam measures are working.  The more
adventurous will have generated statistics to show how many hits each of
their anti-spam measures has had in the last week, and possibly even
generated some graphs to clearly illustrate these statistics to management
or users.\footnote{This was the author's first real foray into processing
Postfix log files.}  Very few will have performed in-depth parsing and
analysis of their log files, where the parsing must correlate the log lines
per-connection or per-queueid rather than processing log lines
independently.  One of the barriers to this kind of processing is the
unstructured nature of Postfix log files, where each log line was added on
an ad hoc basis as a requirement was discovered or new functionality was
added.\footnote{A history of all changes made to Postfix is distributed
with the source code, available from \url{http://www.postfix.org/}} Further
complication arises because the set of rejection messages is not fixed: new
messages can be added by the administrator with custom checks; every
\DNSBL{}\footnote{This document is supplied with a glossary, see
\textsection\ref{Glossary}.} returns a different explanatory message;
policy servers may log different messages depending on the characteristics
of the connection; there are many ways in which the log lines may differ
between servers, even within the same organisation --- servers may be
configured differently, or running different versions of Postfix.  This
paper documents the difficult process of parsing Postfix log files,
presenting \PLP{}, a program that parses log files and places the resulting
data into a database for later use.  The gathered data can then be used to
optimise current anti-spam measures, provide a baseline to test new
anti-spam measures against, or to produce statistics showing how effective
those measures are.  Numerous other uses are possible for such data:
improving server performance by identifying troublesome destinations and
reconfiguring appropriately; identifying regular high volume uses (e.g.\
customer newsletters) and restricting those uses to off-peak times;
detecting virus outbreaks that propagate via email; as a base for billing
customers on a shared server.  Preserving the raw data enables users to
develop a multitude of uses far beyond those conceived of by the author.

\vspace{1em}\noindent\textbf{Layout of the paper:}

Section~\ref{background} provides background information useful in
understanding the paper, parser and algorithm.  It introduces the idea of
using a database schema as an \API{}, providing an interface to the data
gathered that is language-neutral.  The novel separation of rules, actions
and framework is discussed, giving the reasons that approach was taken when
designing the parser.

>>>>

Most mail server administrators will have performed some basic processing
of the log files produced by their mail server at one time or another,
whether it was to debug a problem, explain to a user why their mail is
being rejected, or check if new anti-spam measures are working.  The more
adventurous will have generated statistics to show how successful each of
their anti-spam measures has been in the last week, and possibly even
generated some graphs to clearly illustrate these statistics to management
or users.\footnote{This was the author's first real foray into processing
Postfix log files.}  Very few will have performed in-depth parsing and
analysis of their log files, where the parsing must correlate the log lines
per-connection or per-queueid rather than processing log lines
independently.  One of the barriers to this kind of processing is the
unstructured nature of Postfix log files, where each log line was added on
an ad hoc basis as a requirement was discovered or new functionality was
added.\footnote{A history of all changes made to Postfix is distributed
with the source code, available from \url{http://www.postfix.org/}} Further
complication arises because the set of rejection messages is not fixed: new
messages can be added by the administrator with custom checks; every
\DNSBL{}\footnote{This document is supplied with a glossary, see
\textsection\ref{Glossary}.} returns a different explanatory message;
policy servers may log different messages depending on the characteristics
of the connection; there are many ways in which the log lines may differ
between servers, even within the same organisation --- servers may be
configured differently, or running different versions of Postfix.  This
paper documents the difficult process of parsing Postfix log files,
presenting \PLP{}, a program that parses Postfix log files and places the resulting
data into a database for later use.  The gathered data can then be used to
optimise current anti-spam measures, provide a baseline to test new
anti-spam measures against, or to produce statistics showing how effective
those measures are.  Numerous other uses are possible for such data:
improving server performance by identifying troublesome destinations and
reconfiguring appropriately; identifying regular high volume uses (e.g.\
customer newsletters) and restricting those uses to off-peak times;
detecting virus outbreaks that propagate via email; as a base for billing
customers on a shared server.  Preserving the raw data enables users to
develop a multitude of uses far beyond those conceived of by the author.

\vspace{1em}\noindent\textbf{Layout:}

Section~\ref{background} provides background information useful in
understanding the thesis, parser, and architecture.  It introduces the idea
of using a database schema as an \API{}, providing an interface to the data
gathered that is language-neutral.  The novel separation of rules, actions
and framework is discussed, giving the reasons that approach was taken when
designing the parser.
